### 2018-08-07 paper/finitecomplete-trace.tex
initial files
Theorem on trace refinement testing added

% ==========================================================================
\section{Finite Complete Test Suites for CSP Failures Refinment}
\label{sec:finitecompletefails}
% ==========================================================================

% ==========================================================================

>>>>

% ==========================================================================
\section{Finite Complete Test Suites for CSP Trace Refinement}
\label{sec:finitecomplete}
% ==========================================================================

For establishing trace refinement, the following class of adaptive test cases will be used; again, they are defined for integers $p \ge 0$.

\begin{eqnarray}
U_T(p) & = & U_T(p,\varepsilon)
\\
U_T(p,s) & = & \big( e:(\Sigma - [P/s]^0) \then \efail\then \Stop \big)
\label{eq:ufa}
\\ & & \extchoice \nonumber
\\ & & ([P/s]^0 = \varnothing)    \&   \big( \epass \then \Stop \big)
\label{eq:ufb}
\\ & & \extchoice \nonumber
\\ & & (\#s < p) \& \big( e:[P/s]^0 \then U_T(p,s.e) \big)
\label{eq:ufc}
\\ & & \extchoice \nonumber
\\ & & (\#s = p) \& \big( \epass\then \Stop  \big)
\label{eq:ufd}
\end{eqnarray}

The difference between adaptive tests $U_T(p)$ for trace refinement and
$U_F(p)$ for failures refinement consists in the fact that the former do not ``probe''
the SUT with respect to minimal sets of events to be accepted without blocking.

% ==========================================================================
The existence of complete, finite test suites is expressed in analogy to Theorem~\ref{th:failurestest}. A noteworthy difference consists in the fact that
the complete suite for trace refinement just needs the single adaptive test case 
$U_T(pq-1)$, while the complete failures test suite required the execution of 
$\{ U_F(0),\dots,U_F(pq-1)\}$. The reason for this is that $U_T(pq-1)$ identifies trace
errors for all traces up to length $pq$, while $U_F(pq-1)$ only probes for erroneous
acceptances at the end of each trace of length $(pq -1)$.

\begin{theorem}\label{th:tracetest}
Let $P$ be a divergence-free CSP process over alphabet $\Sigma$ 
whose normalised transition graph $G(P)$ has $p$ states. Define fault domain ${\cal D}$ as
the set of all divergence-free CSP processes over alphabet $\Sigma$, whose transition graph
has at most $q$ states with $q \ge p$. 
Then the test suite 
\[
\TS_T = \{ U_T(pq-1)   \}
\]
is complete with respect to ${\cal F} = (P,\lessdet_T,{\cal D})$.
\xbox
\end{theorem}

We skip the proof of Theorem~\ref{th:tracetest}, since it is just a simplified version of 
the proof of Theorem~\ref{th:failurestest}.

### 2018-08-07 paper/introduction.tex
Theorem on trace refinement testing added

Therefore,  a complementary approach to the definition of CSP fault
domains is presented in this paper. To this end, we observe that every finite-state CSP
process can be semantically represented as a finite normalised transition
graph, whose edges are labelled by the events the process engages in, and
whose nodes are labelled by minimal acceptances or, alternatively, maximal
refusals~\cite{Roscoe:1994:CME:197600}. The maximal refusals  express the
degree of nondeterminism that is present in a given CSP process state that is
in one-one-correspondence to a node of the normalised transition graph.
Inspired by the way that fault-domains are specified for finite state
machines, we define them here as the set of CSP processes whose
normalised transition graphs do not exceed the size of the reference model's
graph by more than a give constant.

Our main contributions in this paper consist in the proof that 
for fault domains of the described type, finite, complete
    test suite generation methods can be given for verifying (1)~trace
    equivalence, (2)~trace refinement, (3)~failures equivalence, and
    (4)~failures refinement.

The existence of -- possibly infinite -- complete test suites has been established for process algebras in general and  for CSP in particular by several authors, 
e.g.~\cite{Hennessy:1988:ATP:50497,Schneider:1995:OST:203471.203475,DBLP:conf/fm/PeleskaS96,peleska1997a,DBLP:conf/pts/CavalcantiS17}. To our best knowledge, however, this article is the first to present {\it finite}, complete test suites associated with this class of fault domains.

>>>>

Therefore,  a complementary approach to the definition of CSP fault
domains is presented in this paper. To this end, we observe that every finite-state CSP
process can be semantically represented as a finite normalised transition
graph, whose edges are labelled by the events the process engages in, and
whose nodes are labelled by minimal acceptances or, alternatively, maximal
refusals~\cite{Roscoe:1994:CME:197600}. The maximal refusals  express the
degree of nondeterminism that is present in a given CSP process state that is
in one-one-correspondence to a node of the normalised transition graph.
Inspired by the way that fault-domains are specified for finite state
machines, we define them here as the set of CSP processes whose
normalised transition graphs do not exceed the size of the reference model's
graph by more than a given constant.

The main contribution of this paper consists in the proof that 
for fault domains of the described type, finite, complete
    test suite generation methods can be given for testing against the conformance relations 
    trace refinement and failures refinement. 
The existence of -- possibly infinite -- complete test suites has been established for process algebras in general and  for CSP in particular by several authors, 
e.g.~\cite{Hennessy:1988:ATP:50497,Schneider:1995:OST:203471.203475,DBLP:conf/fm/PeleskaS96,peleska1997a,DBLP:conf/pts/CavalcantiS17}. To our best knowledge, however, this article is the first to present {\it finite}, complete test suites associated with this class of fault domains and conformance relations.

### 2018-08-07 paper/preliminaries.tex
Theorem on trace refinement testing added

The total function $r$ maps each node $n$ to its \emph{refulsals} $r(n) =
\refs(n)$. Each element of $r(n)$ is a set of events that the CSP process $P$
might refuse to engage into, when in a process state corresponding to $n$.
The number of refusal sets in $\refs(P/s)$ specifies the degree of
nondeterminism that is present in process state $P/s$: the more refusal sets
contained in  $\refs(P/s)$, the more nondeterministic is the behaviour in
state $P/s$. If $P/s$ is deterministic, its refusals coincide with the set of
subsets of $\Sigma - [P/s]^0$, including the empty set.

For finite CSP processes, since the refusals of each process state are
subset-closed~\cite{Hoare:1985:CSP:3921,Roscoe2010}, $\refs(P/s)$ can be
re-constructed by knowing the set of \emph{maximal refusals}
$\maxrefs(P/s)\subseteq\refs(P/s)$. More formally, the maximal refusals
$\maxrefs(P/s)$ are defined as
\begin{equation}
\maxrefs(P/s) = \{ R \in\refs(P/s)~|~\forall R'\in \refs(P/s) - \{ R\}: R \not\subseteq R'\}
\end{equation}
Conversely, with the maximal refusals at hand, we can reconstruct the refusals by subset-closure:
\begin{equation}
\refs(P/s) = \{ R'\in\power(\Sigma)~|~\exists R\in \maxrefs(P/s): R'\subseteq R \}.
\end{equation}
To see that this approach works only for finite CSP processes, consider the
example where $\Sigma$ is infinite. In this case,
$\maxrefs(STOP/\langle\rangle)$ is empty, and so we cannot use this set to
calculate the refusals of $STOP$, that is, $\refs(STOP/\langle\rangle)$ as
defined above. As with refusals, we also use the transition graph-oriented
notation $\maxrefs(n) \subseteq r(n)$ to denote the maximal refusals
associated with graph state $n$: if $n$ is the state reached in the
transition graph by following the edge labels in trace $s$, then $\maxrefs(n)
= \maxrefs(P/s)$.

Well-formed normalised transition graphs must not refuse an event of the
fan-out of a state in {\it every} refusal applicable in this state; more
formally,
\begin{equation}
\label{eq:wellformedg}
\forall n\in N, e\in\Sigma: (n,e)\in\dom~t \Rightarrow
\exists R\in \maxrefs(n): e\not\in R
\end{equation}

@@

Every node of a transition graph can be labelled with with its minimal acceptances, and this information is equivalent to 
that contained in its maximal refusals.

@@

Note that the node names are generated by the FDR tool (see next paragraph).
The node numbering is generated by FDR during the normalisation procedure.
Therefore, the node numbers do not reflect the distance from the initial node
Node\_0.
%Node\_2 is a direct successor to the initial node Node\_0, while Node\_1 is a direct successor of Node\_2.
\end{example}

>>>>

The total function $r$ maps each node $n$ to its \emph{refulsals} $r(n) =
\refs(n)$. Each element of $r(n)$ is a set of events that the CSP process $P$
might refuse to engage into, when in a process state corresponding to $n$.
 If $P/s$ is deterministic, its refusals coincide with the set of
subsets of $\Sigma - [P/s]^0$, including the empty set.

For finite CSP processes, since the refusals of each process state are
subset-closed~\cite{Hoare:1985:CSP:3921,Roscoe2010}, $\refs(P/s)$ can be
re-constructed by knowing the set of \emph{maximal refusals}
$\maxrefs(P/s)\subseteq\refs(P/s)$. More formally, the maximal refusals
$\maxrefs(P/s)$ are defined as
\begin{equation}
\maxrefs(P/s) = \{ R \in\refs(P/s)~|~\forall R'\in \refs(P/s) - \{ R\}: R \not\subseteq R'\}
\end{equation}
Conversely, with the maximal refusals at hand, we can reconstruct the refusals by subset-closure:
\begin{equation}
\refs(P/s) = \{ R'\in\power(\Sigma)~|~\exists R\in \maxrefs(P/s): R'\subseteq R \}.
\end{equation}
The cardinality of $\maxrefs(P/s)$ reflects the degree of
nondeterminism that is present in process state $P/s$: the more maximal refusal sets
contained in  $\maxrefs(P/s)$, the more nondeterministic is the behaviour in
state $P/s$. Deterministic process states $P/s$ have exactly the one maximal refusal
$\Sigma-[P/s]^0$.

To see that this approach works only for finite CSP processes, consider the
example where $\Sigma$ is infinite. In this case,
$\maxrefs(\Stop/\langle\rangle)$ is empty, and so we cannot use this set to
calculate the refusals of $\Stop$, that is, $\refs(\Stop/\langle\rangle)$ as
defined above. As with refusals, we also use the transition graph-oriented
notation $\maxrefs(n) \subseteq r(n)$ to denote the maximal refusals
associated with graph state $n$: if $n$ is the state reached in the
transition graph by following the edge labels in trace $s$, then $\maxrefs(n)
= \maxrefs(P/s)$.

Well-formed normalised transition graphs must not refuse an event of the
initials of a state in {\it every} refusal applicable in this state; more
formally,
\begin{equation}
\label{eq:wellformedg}
\forall n\in N, e\in\Sigma: (n,e)\in\dom~t \Rightarrow
\exists R\in \maxrefs(n): e\not\in R
\end{equation}

@@

Every node of a transition graph can alternatively 
be labelled  with its minimal acceptances, and this information is equivalent to 
that contained in its maximal refusals.

@@

Note that the node names including their number suffixes 
are generated by the FDR tool (see next paragraph).
The numbering is generated   during the normalisation procedure.
Therefore, the node numbers do not reflect the distance from the initial node
Node\_0.
\end{example}
